---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

This assignment was accomplished by first forking rdpeng's [repository.](https://github.com/rdpeng/RepData_PeerAssessment1)

The data for the assignment is included in the repository in a file named **activity.zip**

Fork the assignment repository above and clone to your desired working directory.

## Loading and preprocessing the data

Load the required libraries

```{r}
library(dplyr)
library(lubridate)
library(ggplot2)
```

Load and Process Data

```{r}
if (!file.exists("data")) {
        dir.create("data")
        }
unzip("./activity.zip", exdir = "./data")
data <- read.csv("./data/activity.csv")
data[,2] <- as_date(data[,2])
data[,1] <- as.numeric(data[,1])
```

## What is mean total number of steps taken per day?

Calculate the total number of steps taken per day

```{r}
spd <- data %>% 
        group_by(date) %>%
        summarize(steps = sum(steps)) %>%
        as.data.frame(spd)
spd
```

Make a histogram of the total number of steps taken each day

```{r}
hist(spd$steps, main = "Total Steps Per Day", xlab = "Steps", col = "red", breaks = 10)
rug(spd$steps)
```

Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean(spd$steps, na.rm = TRUE)
median(spd$steps, na.rm = TRUE)
```

## What is the average daily activity pattern?

Make a time series plot (i.e. type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
ints <- data %>%
        group_by(interval) %>%
        summarise(steps = mean(steps, na.rm = TRUE)) %>%
        as.data.frame(ints)
plot(x = ints$interval, y = ints$steps, type = "l", xlab = "5-minute Interval of Day",
        ylab = "Average Number of Steps (All Days)", main = 
        "Average Number of Steps per Interval Across All Days")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
filter(ints, steps == max(ints[,2]))
```

## Imputing missing values

Calculate and report the total number of missing values in the dataset

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Create a new dataset that is equal to the original dataset but with the missing data filled in.

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

Do these values differ from the estimates from the first part of the assignment?

What is the impact of imputing missing data on the estimates of the total daily number of steps?

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

Make a panel plot containing a time series plot (i.e. \color{red}{\verb|type = "l"|}type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
