---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

This assignment was accomplished by first forking rdpeng's [repository.](https://github.com/rdpeng/RepData_PeerAssessment1)

The data for the assignment is included in the repository in a file named **activity.zip**

Fork the assignment repository above and clone to your desired working directory.

## Loading and preprocessing the data

Load the required libraries

```{r}
library(dplyr)
library(lubridate)
library(ggplot2)
```

Load and Process Data

```{r}
if (!file.exists("data")) {
        dir.create("data")
        }
unzip("./activity.zip", exdir = "./data")
data <- read.csv("./data/activity.csv")
data[,2] <- as_date(data[,2])
data[,1] <- as.numeric(data[,1])
```

## What is mean total number of steps taken per day?

Calculate the total number of steps taken per day

```{r}
spd <- data %>% 
        group_by(date) %>%
        summarize(steps = sum(steps)) %>%
        as.data.frame(spd)
spd
```

Make a histogram of the total number of steps taken each day

```{r, fig.width = 6, fig.height = 6}
hist(spd$steps, main = "Total Steps Per Day", xlab = "Steps", col = "red", breaks = 10)
rug(spd$steps)
```

Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean(spd$steps, na.rm = TRUE)
median(spd$steps, na.rm = TRUE)
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
